---
title: "Image preprocessing"
author: "Ian"
date: "2023-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (Will explore more and update this file)

## Load libraries
```{r, message=FALSE, warning=FALSE}
library(EBImage)
library(imager)
library(SpatialPack)

###If you haven't install
#install.packages("BiocManager")
#BiocManager::install("EBImage")

#install.packages(c("imager"),("SpatialPack"))
```

## Load image
```{r}
img = load.image("data/Biotechnology_small/data_processed/cell_images/cluster_1/cell_4759.png")
img = resize(img,50,50)
EBImage::display(img, method = "raster")
```

Filter-library(SpatialPack): https://search.r-project.org/CRAN/refmans/SpatialPack/html/denoise.html
not sure if this one is that useful.
```{r}
img_denoise = denoise(as.matrix(img), type="Lee", looks = 0.05, damping = 0.1)
par(mfrow=c(1,2))
EBImage::display(img, method = "raster")
EBImage::display(img_denoise, method = "raster")
```

Power law
what is power law: 
https://www.tutorialspoint.com/dip/gray_level_transformations.htm#:~:text=Power%20%E2%80%93%20Law%20transformations&text=This%20type%20of%20transformation%20is,displayed%20on%20CRT%20is%20dark.
```{r}
gamma <- 2.5
img_transformed <- 255 * (img / 255) ^ gamma 
par(mfrow=c(1,2))
plot(img, main="Original Image")
plot(img_transformed, main="Transformed Image")
```


EBImage
Thresholding + Opening
Opening(wikipedia): removes small objects from the foreground (usually taken as the bright pixels) of an image, placing them in the background
```{r}
###If the image is 50*50 use the 3*3 window, but if we use the original image we will need a bigger window size, like w=10,h=10
#nmask = thresh(img, w=10, h=10, offset=0.05)
nmask = thresh(img, w=3, h=3, offset=0.05)
nmask = opening(nmask, makeBrush(3, shape='diamond'))

par(mfrow=c(1,2))
plot(img, main="Original Image")
plot(nmask, main="nmask")
```

## Resources

Video - Bioimage Analysis 2: Pre-Processing (How people deal with cell images): 
https://www.youtube.com/watch?v=MIDK8BqJY_8

Video - Bioimage Analysis 3: Segmentation (Thresholding, outline): 
https://www.youtube.com/watch?v=jLd2I2adQtw

Introduction to EBImage:  
https://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html#1_Getting_started


