---
title: "Research Question"
subtitle: "DATA3888 Data Science Capstone"
output:
  html_document:
    theme: cerulean
    code_folding: hide 

date: "2023-05-28"
author: "Biotechnolgoy and Platforms - Image07 "
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Executive Summary 

**What is the problem?**

**The main findings**

**Key figures if needed**

**Practical relevance of the study**


## Background 

**ensure understanding of biotechnological data and platforms** 

## Aims

## Method 

**Add innovations for each aspect**
**ensure understanding of biotechnological data and platforms**
**can mention how rd pixels was added but scrapped and also how resnet was almost added but scrapped**
**add how each part is reproducible if neede**

### Data collection 
### Developed Models
#### AlexNet
#### Random Forest 
### Evaluation Strategies 
#### Accuracy
#### Robustness
#### Interpretability 
#### qualitative to ensure communication 


## Results - Research Question 

### Performance 

#### Accuracy
#### Robustness
**also mention generality of findings**

```{r}
# this is where all the code will be added --
# for EVERYTHING 
```



### Interpretability 


## Results - Product 

### Clear description and illustration 

### concepts from multiple disciplines 

**add in the video maybe**

### evaluation metric for the app - feedback and interview session 

## Discussion 

### limitations 
### issues with development process
### issues with product

### Future work

### improvemtns in both discipline 

## Conclusion 



## Student Contributions 

**Akasha**
**Thoon**
**Ian**
**James**
**George**
**Nikki**

## Appendix 
